### 2022-08-22 doc/quickstart.tex
Recorded a latex screencast

\documentclass{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{mdframed}
\usepackage{minted}
\renewcommand{\MintedPygmentize}{/nix/store/1r3k7gbfc8bjc4mr5f3bc3j7g7ws24cy-python3-3.8.12-env/bin/pygmentize}
\usepackage{verbatim}

\newenvironment{myverbatim}
  {\VerbatimEnvironment
   \begin{minted}[breaklines,fontsize=\footnotesize]{latex}}
  {\end{minted}}
\BeforeBeginEnvironment{myverbatim}{\begin{mdframed}[nobreak=true,frametitle=\tiny{Source}]}
\AfterEndEnvironment{myverbatim}{\end{mdframed}}

\newenvironment{lcode}
  {\VerbatimEnvironment
   \begin{minted}[breaklines,fontsize=\footnotesize]{python}}
  {\end{minted}}
\BeforeBeginEnvironment{lcode}{\begin{mdframed}[frametitle=\tiny{Rendered}]}
\AfterEndEnvironment{lcode}{\end{mdframed}}

\newenvironment{lresult}{\verbatim}{\endverbatim}
\BeforeBeginEnvironment{lresult}{\begin{mdframed}\footnotesize}
\AfterEndEnvironment{lresult}{\end{mdframed}}

\newcommand{\linline}[2]{#2}

\graphicspath{ {./} }

\begin{document}

\section{Organizing the document}

\subsection{Preamble}

The following Latex code provides a reference Latex definitions of
\texttt{lcode}, \texttt{lresult} and \texttt{linline} tags. These names are
currently hardcoded into VimREPL. Users may define them arbitrary.

\begin{myverbatim}
\newenvironment{lcode}
  {\VerbatimEnvironment
   \begin{minted}[breaklines,fontsize=\footnotesize]{python}}
  {\end{minted}}
\BeforeBeginEnvironment{lcode}{\begin{mdframed}[frametitle=\tiny{Rendered}]}
\AfterEndEnvironment{lcode}{\end{mdframed}}

\newenvironment{lresult}{\verbatim}{\endverbatim}
\BeforeBeginEnvironment{lresult}{\begin{mdframed}\footnotesize}
\AfterEndEnvironment{lresult}{\end{mdframed}}

\newcommand{\linline}[2]{#2}
\end{myverbatim}

\subsection{Code and result sections}

%lignore
\begin{myverbatim}
\begin{lcode}
W='Value of \\textbf{W}!'
print(W)
\end{lcode}

\begin{lresult}
\end{lresult}
\end{myverbatim}
%lnoignore

\begin{lcode}
W='Value of \\textbf{W}!'
print(W)
\end{lcode}

\begin{lresult}
Value of \textbf{W}!
\end{lresult}

\subsection{Verbatim input and output}

\subsubsection{Verbatim output}

%lignore
\begin{myverbatim}
\begin{lcode}
print(W)
\end{lcode}
\begin{mdframed}
%lresult
%lnoresult
\end{mdframed}
\end{myverbatim}
%lnoignore

\begin{lcode}
W='Value of \\textbf{W}!'
print(W)
\end{lcode}
\begin{mdframed}
%lresult
Value of \textbf{W}!
%lnoresult
\end{mdframed}

\subsubsection{Verbatim input}
A less often used option of verbatim input is also supported

%lignore
\begin{myverbatim}
%lcode
print('Value of \\textbf{W}!')
%lnocode
\begin{lresult}
\end{lresult}
\end{myverbatim}
%lnoignore

%lcode
print(r'Value of \textbf{W}!')
%lnocode
\begin{lresult}
Value of \textbf{W}!
\end{lresult}

\subsection{Inline code}

%lignore
\begin{myverbatim}
\begin{lcode}
W='\\textbf{W}'
\end{lcode}
\begin{mdframed}Here is an inline value of \linline{W}{}.\end{mdframed}
\end{myverbatim}
%lnoignore

\begin{lcode}
W='\\textbf{W}'
\end{lcode}
\begin{mdframed}Here is an inline value of \linline{W}{\textbf{W}}.\end{mdframed}

\break
\section{Mathplotlib and other graphics}

%lignore
\begin{myverbatim}
\begin{lcode}
import os
import numpy as np
import matplotlib.pyplot as plt
plt.close()
X=np.array(range(100))
plt.plot(X, np.sin(X/10))
os.makedirs('img',exist_ok=True)
path='img/plot.png'
plt.savefig(path)
\end{lcode}

\begin{center}
\includegraphics[width=0.8\textwidth]{\linline{path}{img/plot.png}}
\end{center}
\end{myverbatim}
%lnoignore

\begin{lcode}
import os
import numpy as np
import matplotlib.pyplot as plt
plt.close()
X=np.array(range(100))
plt.plot(X, np.sin(X/10))
os.makedirs('img',exist_ok=True)
path='img/plot.png'
plt.savefig(path)
\end{lcode}

\begin{center}
\includegraphics[width=0.8\textwidth]{\linline{path}{img/plot.png}}
\end{center}

\section{Live Demo}

\begin{lcode}
W='Hello'
print(W)
\end{lcode}

\begin{lresult}
Hello
\end{lresult}

\end{document}